### 2024-02-20 code/processing-code/processingfile-v1.qmd
Part 2 Updates
We fixed an error in one of the variables we created, total_larcster. We began our EDA with histograms and scatter plots with numeric variables.

---
title: "Processing Script"
author: "Taylor Glass and Arlyn Santiago"
date: "2024-02-16"
output: html_document
---

# Processing script

# Setup

Load the necessary packages for data loading and processing.

```{r}
#| message: false
library(readr) # for loading csv files
library(readxl) # for loading excel files (codebook)
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(naniar) #to explore data missingness
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(tidymodels) #to impute missing data
library(recipes) #to impute missing data
```

# Data loading

```{r}
#| message: false
# path to data using here() function
data_location <- here("data","raw-data","AIU All Women Dataset.csv")
rawdata <- read_csv(data_location)
```

# Check data

We will load and examine the codebook first.

```{r}
#| message: false
# write file path to the codebook
codebook_location <- here("data", "AIU All Women Dataset Codebook.xlsx")

# specify the 3 sheets of the codebook we need 
sheets <- c("Demographics", "Contraceptive need & use", "Pregnancy-related & newborn")

# load codebook
codebook <- lapply(sheets, function(sheet) {
  readxl::read_excel(codebook_location, sheet = sheet)
})
print(codebook)
```
Note: should we include other covariates like marriage percentages, number of live births, number of induced abortions

# Glimpse at the data

We will take a glimpse at our data to better understand the variables we are working with. There are four character variables: `country`, `data_source_year`, `region`, and `subregion`. The remainder of the variables are numeric and describe percentage of birth outcomes, maternal deaths due to birth outcomes, contraceptive methods divided by age groups, wealth levels in relation to contraceptive methods, various abortion procedure rates, current cost of pregnancy outcomes and STI treatment.

```{r}
#| message: false
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```
# Identifying variables to be cleaned 
We want to explore the relationship between unintended pregnancies and birth control methods across different age groups. We will use three demographic variable: `country`, `region`, and `subregion`. The variable for the total percentage of unintended pregnancies, `pct_upreg` will be the main outcome variable. We could also use `upreg_subreg` to explore unintended pregnancies by subregion. The predictor variables will include percentage of unintended pregnancies across three types of birth control: long acting reversible methods (`pct_upreg_larcster`), short term reversible methods (`pct_upreg_sarc`), traditional methods (`pct_upreg_trad`), and no method (`pct_upreg_nouse`). No method variables can serve as control groups for statistical tests. Covariates will include the number of women using each of these methods across four different age groups: 15-19, 20-24, 25-34, and 35-49. 
-Number of women 15-19 using long-acting reversible methods `larc_1519`
-Number of women 15-19 using short-acting methods `sarc_1519`
-Number of women 15-19 using traditional methods `tradmethods_1519`
-Number of women 15-19 using no method `nomethod_1519`
-Number of women 20-24 using long-acting reversible methods `larc_2024`
-Number of women 20-24 using short-acting methods	`sarc_2024`
-Number of women 20-24 using traditional methods	`tradmethods_2024`
-Number of women 20-24 using no method	`nomethod_2024`
-Number of women 25-34 using long-acting reversible methods	`larc_2534`
-Number of women 25-34 using short-acting methods	`sarc_2534`
-Number of women 25-34 using traditional methods	`tradmethods_2534`
-Number of women 25-34 using no method	`nomethod_2534`
-Number of women 35-49 using long-acting reversible method	`larc_3549`
-Number of women 35-49 using short-acting method	`sarc_3549`
-Number of women 35-49 using all traditional methods	`tradmethods_3549`
-Number of women 35-49 using no method	`nomethod_3549`

code for low, medium, and high use per country using case_when()

```{r}
#| message: false
# select the outcome variable, predictor variables, and covariates  
rawdata2 <- rawdata %>% 
                select(country, region, subregion, pct_upreg, upreg_subreg, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad, pct_upreg_nouse, larc_1519, ster_1519, sarc_1519, tradmethods_1519, nomethod_1519, larc_2024, ster_2024, sarc_2024, tradmethods_2024, nomethod_2024, larc_2534, ster_2534, sarc_2534, tradmethods_2534,  nomethod_2534, larc_3549, ster_3549, sarc_3549, tradmethods_3549, nomethod_3549)

# take a glimpse at the selected variables 
skim(rawdata2)
head(rawdata2)
```

# Cleaning

By inspecting the data as done above, we find some problems that need addressing:

First, the `region` variable is a character variable, but it will be more useful later as a factor. We can easily mutate it from a character variable to factor variable. All the remaining predictor and covariate variables are numeric as they should be. 
```{r}
#| message: false
# convert character variable to factor variable
rawdata2 <- rawdata2 %>% 
              mutate(region = as.factor(region))

# check that the conversion worked 
class(rawdata2$region)
```

Second, there are 24 entries missing for the percentage of all unintended pregnancies by country `pct_upreg`, which accounts for 18.18% of the observations. After isolating the observations with this missing variable, we see that the majority of countries in this category are located in Asia. This does not make sense because the majority of countries in the raw dataset are located in Africa, which suggests that the data are not missing at random. 
```{r}
#| message: false
# find missing percentage for each variable
miss_var_summary(rawdata2)

# isolate the percentage of all unintended pregnancies by country to discover trends in missingness 
variable_with_missing <- "pct_upreg"
missing_obs <- rawdata2[is.na(rawdata2[[variable_with_missing]]), ]
other_variables <- colnames(rawdata2)[!colnames(rawdata2) == variable_with_missing]
pct_upregmissing <- missing_obs[, other_variables]
print(pct_upregmissing)

# compare dataset with the missing variable to the full dataset
table(pct_upregmissing$region)
table(rawdata2$region)
```

We should impute the missing data here to avoid introducing bias because 18.18% of the data is missing, and this variable is essential for answering the research question. After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data.  
```{r}
#| message: false
# find mean to use for imputed data
summary(rawdata2$pct_upreg)

# create a recipe to impute data 
rec <- recipe(pct_upreg ~ ., data = rawdata2) %>%
  step_naomit(all_predictors(), -all_outcomes(), skip = TRUE) %>%
  step_impute_mean(all_predictors()) #which predictors would I use?

# prepare the raw data using the recipe
data_prep <- prep(rec, training = rawdata2)

# impute the data 
data_imputed <- bake(data_prep, new_data = NULL) #all pct_upreg are NaN values

# identify rows with missing values in the original dataset
missing_rows <- which(is.na(rawdata2$pct_upreg))

# replace missing values in the original dataset with imputed values
rawdata2$pct_upreg[missing_rows] <- data_imputed$pct_upreg[missing_rows]

```
Once the missing values have been imputed, we should further clean the `pct_upreg` variable to have a typical percentage format. 
```{r}
#| message: false
# convert percentage variable to typical percentage format
rawdata2 <- rawdata2 %>% 
              mutate(pct_upreg = pct_upreg * 100)

# ensure mutation worked correctly
summary(rawdata2$pct_upreg)
```

We also need to evaluate the 108 missing entries of the percentage of unintended pregnancies by subregion `upreg_subreg`. Since 81.81% of this variable is missing, it does not make sense to include it in the analysis because we do not have enough information to impute the missing data. We will remove this variable for the sake of this analysis.
```{r}
#| message: false
# isolate the percentage of all unintended pregnancies by subregion to discover trends in missingness
variable_with_missing <- "upreg_subreg"
missing_obs2 <- rawdata2[is.na(rawdata2[[variable_with_missing]]), ]
other_variables2 <- colnames(rawdata2)[!colnames(rawdata2) == variable_with_missing]
pct_subregmissing <- missing_obs2[, other_variables2]
print(pct_subregmissing)

# remove unintended pregnancies by subregion variable due to missingness
rawdata2 <- rawdata2 %>% 
              select(-upreg_subreg)
```

Fortunately, the rest of the variables required for this analysis are complete. Each of the four variables describing the percentage of unintended pregnancies based on birth control type also need to be converted to traditional percentage format. 
```{r}
#| message: false
# convert percentage variable to typical percentage format
rawdata2 <- rawdata2 %>% 
              mutate(pct_upreg_larcster = pct_upreg_larcster * 100, 
                     pct_upreg_sarc = pct_upreg_sarc * 100, 
                     pct_upreg_trad = pct_upreg_trad * 100, 
                     pct_upreg_nouse = pct_upreg_nouse * 100)

# ensure mutations worked correctly
summary(rawdata2$pct_upreg_larcster)
summary(rawdata2$pct_upreg_sarc)
summary(rawdata2$pct_upreg_trad)
summary(rawdata2$pct_upreg_nouse)
```

To make the covariate variables more useful, it makes sense to create a new factor variable denoting low, medium, and high usage of each birth control method by country. First, we need to determine the boundaries for each level of usage by exploring how many women use each method across age groups. We need to create a variable showing the total usage of each birth control method to remove the effect of age. It is worth noting here that the long acting reversible and sterilzation methods are combined in the`pct_upreg_larcster` variable, so we need to consider all the women using both of these methods in the total. 
```{r}
#| message: false
# create new variables showing total usage of each method including all age groups
rawdata2 <- rawdata2 %>% 
              mutate(total_larcster = larc_1519 + larc_2024 + larc_2534 + larc_3549 + ster_1519 + ster_2024 + ster_2534 +                          ster_3549,
                     total_sarc = sarc_1519 + sarc_2024 + sarc_2534 + sarc_3549,
                     total_trad = tradmethods_1519 + tradmethods_2024 + tradmethods_2534 + tradmethods_3549,
                     total_none = nomethod_1519 + nomethod_2024 + nomethod_2534 + nomethod_3549)

# explore summary of each total to set boundaries for each level of usage
rawdata2 %>% skim(total_larcster, total_sarc, total_trad, total_none)
```
After finding the distribution of number of women using each type of birth control, we can use that information to create a new factor variable denoting low, medium, and high usage for each method. The "low" category will include observations below the 25th percentile. The "medium" category will include observations between the 25th and 75th percentiles. The "high" category will include observations above the 75th percentile. 
```{r}
#| message: false
# create usage variable for long acting reversible methods 
rawdata2 <- rawdata2 %>% 
              mutate(larcster_usage = as.factor(case_when(
                                  total_larcster < 23.94865 ~ "Low",
                                  total_larcster >= 23.94865 & total_larcster < 1554.7609 ~ "Medium",
                                  total_larcster >= 1554.7609 ~ "High")))

# check if new variable was created correctly 
class(rawdata2$larcster_usage)

# create usage variable for short acting reversible methods 
rawdata2 <- rawdata2 %>% 
              mutate(sarc_usage = as.factor(case_when(
                                  total_sarc < 114.33113 ~ "Low",
                                  total_sarc >= 114.33113 & total_sarc < 1554.7609 ~ "Medium",
                                  total_sarc >= 1554.7609	 ~ "High")))

# check if new variable was created correctly 
class(rawdata2$sarc_usage)

# create usage variable for traditional methods 
rawdata2 <- rawdata2 %>% 
              mutate(trad_usage = as.factor(case_when(
                                  total_trad < 15.05000	 ~ "Low",
                                  total_trad >= 15.05000 & total_trad < 332.4750 ~ "Medium",
                                  total_trad >= 332.4750	 ~ "High")))

# check if new variable was created correctly 
class(rawdata2$trad_usage)

# create usage variable for no methods
rawdata2 <- rawdata2 %>% 
              mutate(none_usage = as.factor(case_when(
                                  total_none < 80.07500	 ~ "Low",
                                  total_none >= 80.07500 & total_none < 875.2750 ~ "Medium",
                                  total_none >= 875.2750	 ~ "High")))

# check if new variable was created correctly 
class(rawdata2$none_usage)
```

Lastly, we will assign the clean data to a final variable, which makes it easier to add further cleaning steps above if necessary.

```{r}
#| message: false
processeddata <- rawdata2
```

# Save data

Finally, we will save the clean data as RDS file to preserve coding like factors, characters, numeric, etc.

```{r}
#| message: false
save_data_location <- here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```