### 2024-03-13 code/eda-code/eda.qmd
Additional Predictors
I used the codebook to find 5 more predictors for our outcome. I processed them and made sure there was no missingness. I did a little bit of EDA and created simple linear models with each of them.

---
title: "Exploratory Analysis Script"
author: "Taylor Glass and Arlyn Santiago"
date: "02/21/24"
output: html_document
---

# Setup

```{r}
#load necessary packages
library(here) #for data loading/saving
library(dplyr) # for basic piping functions
library(skimr) # for summarizing data
library(ggplot2) # for visualizing data
library(gtsummary) # for creating tables
```

# Load the data

```{r}
#path to data
data_location <- here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables
We will begin with creating a summary table of our cleaned variables using the skim()function. This provides information about each of the numeric variables including the mean, standard deviation, quartiles, missing observations, and a histogram.

```{r}
# skim the data to create summary file
summary_df <- skim(mydata)
print(summary_df)
# save to file
summarytable_file <- here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

We also created a simple table with the categorical variables: `region`, `larcster_usage`, `sarc_usage`, and `none_usage`.
```{r}
# isolate categorical variables into their own data set
mydata_cat <- mydata %>% select(region, larcster_usage, sarc_usage, trad_usage, none_usage)

# create a summary table with counts of each level of each categorical variable
summary_table <- tbl_summary(mydata_cat, statistic = all_categorical() ~ "{n}")
print(summary_table)

# save to file 
summarytable_file <- here("results", "tables", "summarytablecat.rds")
saveRDS(summary_table, file = summarytable_file)
```

# Data exploration through figures

We will make histogram plots for the continuous outcomes in our dataset.

It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well. We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.

```{r}
# explore the percentage of unintended pregnancies related to long acting reversible methods in each country
p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + 
  geom_histogram() +
  ggtitle("Percent Unintended Pregnancy Using Long Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(p2)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_distribution.png")
ggsave(filename = figure_file, plot=p2)

# explore the percentage of unintended pregnancies related to short acting reversible methods in each country
p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + 
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using Short Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(p3)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png")
ggsave(filename = figure_file, plot=p3)

# explore the percentage of unintended pregnancies related to traditional methods in each country
p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + 
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using Traditional Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(p4)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_distribution.png")
ggsave(filename = figure_file, plot=p4)

# explore the percentage of unintended pregnancies related to no birth control usage in each country
p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) +
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using No Birth Control") + xlab("Percentage") + ylab("Number of countries")
plot(p5)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_nouse_distribution.png")
ggsave(filename = figure_file, plot=p5) 
```
P2: This histogram is highly skewed right, which shows that most countries experience a very low percentage of unintended pregnancies related to long acting reversible and sterilization methods.

P3: This histogram shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods. About half of the countries in the dataset report less than 20% of pregnancies are unintended when women use this form of birth control.

P4: This histogram shows that about half of the countries reported less than 10% percent of unintended pregnancies associated with using traditional birth control methods

P5: This histogram shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.

We will now explore the new total usage variables that were created using the number of women using each method across the four age categories. This set of plots revealed large outliers in the 'total_larcster`, 'total_sarc', 'total_trad', and `total_none` variables, which prevented the graphs from being very useful. The histograms look like most observations for each of these four variables are 0, but that is not possible because the summary table showed that the minimum value for all four of these variables is greater than 0. This distortion is caused by the two large outliers in the data set.

```{r}
# explore the total number of people using long acting reversible methods in each country
p6 <- mydata %>% ggplot(aes(x=total_larcster)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Long Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p6)
figure_file = here("results", "figures", "exploratory-figures", "total_larcster_distribution.png")
ggsave(filename = figure_file, plot=p6)

# explore the total number of people using short acting reversible methods in each country
p7 <- mydata %>% ggplot(aes(x=total_sarc)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Short Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p7)
figure_file = here("results", "figures", "exploratory-figures", "total_sarc_distribution.png")
ggsave(filename = figure_file, plot=p7)

# explore the total number of people using traditional methods in each country
p8 <- mydata %>% ggplot(aes(x=total_trad)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Traditional Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p8)
figure_file = here("results", "figures", "exploratory-figures", "total_trad_distribution.png")
ggsave(filename = figure_file, plot=p8)

# explore the total number of people not using birth control in each country
p9 <- mydata %>% ggplot(aes(x=total_none)) + 
  geom_histogram() + 
  ggtitle("Total Usage of No Birth Control Method per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p9)
figure_file = here("results", "figures", "exploratory-figures", "total_none_distribution.png")
ggsave(filename = figure_file, plot=p9) 
```

The outliers were explored and addressed in the processing-file.qmd in the code > processing-code folder. Now that we have standardized population variables, We will start exploring how many women use each type of birth control in each country compared to the percent of unintended pregnancies related to that birth control. Plot 10 shows `pct_upreg_larcster` vs `larcster_standard` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods. Plot 11 uses `pct_upreg_sarc` and `sarc_standard` to examine the same relationship as plot 10 but with short-acting reversible methods. Plot 12 shows `pct_upreg_trad` versus `trad_standard`, and plot 13 shows `pct_upreg_nouse` versus `none_standard`. 
```{r}
# explore percentage of unintended pregnancies compared to standardized count of women using long-acting reversible and sterilization methods
p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p10)
figure_file = here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png")
ggsave(filename = figure_file, plot=p10)

# explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p11)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
ggsave(filename = figure_file, plot=p11)

# explore percentage of unintended pregnancies compared to standardized count of women using traditional methods
p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p12)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png")
ggsave(filename = figure_file, plot=p12)

# explore percentage of unintended pregnancies compared to standardized count of women not using no method
p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p13)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png")
ggsave(filename = figure_file, plot=p13) 
```

P10: As the standardized count of women using LARCs and sterilization methods increases, the percentage of unintended pregnancies tends to increase. There is minimal variation around the linear equation, excluding the two largest percentages accounted for by India and China.

P11: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARMs, but there is considerably more variation in this relationship.

P12: As percentage of unintended pregnancies associated with traditional methods increases, the total number of women using this method increases. There is minimal variation in the trend when percentage of unintended pregnancies is less than 8%, but there is more variation when the percentage is greater than 10%.

P13: The same expected positive trend is observed for women who are not using a birth control method, but there is considerable variation in this graph across all percentages of unintended pregnancies. Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.

We are also interested in how these relationships between percent of unintended pregnancies using each birth control method and the number of people using each method varies across region. We adjusted the original scatter plots created above by coloring based on region.

```{r}
# explore percent unintended pregnancies using long acting reversible and sterilziation methods by region
p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + 
  geom_point() + geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p14)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png")
ggsave(filename = figure_file, plot=p14)

# explore percent unintended pregnancies using short acting reversible methods by region
p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p15)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
ggsave(filename = figure_file, plot=p15)

# explore percent unintended pregnancies using traditional methods by region
p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') +  
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p16)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png")
ggsave(filename = figure_file, plot=p16)

# explore percent unintended pregnancies using no method by region
p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') +   
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p17)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png")
ggsave(filename = figure_file, plot=p17) 
```
P14: Among women using long acting reversible and sterilization methods, Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.

P15: This graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.

P16: Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.

P17: This graph is much different than the other three, which is to be expected at this point. Oceania has the highest percentage of unintended pregnancies when women are not using birth control. The slope of the relationship for each region appears to be similar in this graph with the exception of Oceania having a steeper slope. There are considerably more outliers in this graph with African countries

Our next group of plots will be violin plots to explore the usage level of each birth control method, a categorical variable, associated with percentage of unintended pregnancies use for each birth control method, a numerical variable.
```{r}
# explore long acting reversible and sterilization method usage compared to percent of unintended pregnancies
p18 <- mydata %>% ggplot(aes(x = larcster_usage, y = pct_upreg_larcster)) +
  geom_violin() +
  ggtitle("Usage of LARCs and Sterilization vs Percent of Unintended Pregnancies") +
  labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
plot(p18)
figure_file = here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p18)

# explore short acting reversible method usage compared to percent of unintended pregnancies
p19 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
  geom_violin() + ggtitle("Usage of SARCs vs Percent of Unintended Pregnancies") +
  labs(x = "Short Acting Reversible Usage", y = "Percentage Unintended Pregnancies") 
plot(p19)
figure_file = here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p19)

# explore traditional method usage compared to percent of unintended pregnancies
p20 <- mydata %>% ggplot(aes(x = trad_usage, y = pct_upreg_trad)) +
  geom_violin() + 
  ggtitle("Usage of Traditional Method vs Percent of Unintended Pregnancies") +
  labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
plot(p20)
figure_file = here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p20)

# explore no method usage compared to percent of unintended pregnancies
p21 <- mydata %>% ggplot(aes(x = none_usage, y = pct_upreg_nouse)) +
  geom_violin() +  
  ggtitle("No Birth Control Usage vs Percent of Unintended Pregnancies") +
  labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
plot(p21)
figure_file = here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p21) 
```
P18: The distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods. High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.

P19: It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.

P20: Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.

P21: While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.

We also need to explore the remaining predictor variables: `region`, `pct_currentlymarried`, `pct_nevermarried`, `rate_matdeaths`, `rate_matdeaths_upreg`, and `rate_matdeaths_abs`. Simple histograms will allow us to see the distribution of each variable and ensure there are no outliers. 
```{r}
# explore the region predictor
p22 <- mydata %>% ggplot(aes(x=region)) + 
  geom_bar() +
  ggtitle("Countries in Each Region") + 
  xlab("Region") + ylab("Number of Countries")
plot(p22)
figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
ggsave(filename = figure_file, plot=p22)

# explore the current marriage predictor
p23 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Currently Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(p23)
figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
ggsave(filename = figure_file, plot=p23)

# explore the never married predictor
p24 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Never Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(p24)
figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
ggsave(filename = figure_file, plot=p24)

# explore the total maternal mortality rate predictor
p25 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
  geom_histogram() +
  ggtitle("Total Rate of Maternal Deaths") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(p25)
figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
ggsave(filename = figure_file, plot=p25)

# explore the maternal mortality rate from unintended pregnancies predictor
p26 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
  geom_histogram() +
  ggtitle("Rate of Maternal Deaths from Unintended Pregnancies") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(p26)
figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
ggsave(filename = figure_file, plot=p26)

# explore outliers in maternal mortality from unintended pregnancies
mydata %>% filter(rate_matdeaths_upreg > 6000)

# explore the maternal mortality rate from abortions
p27 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
  geom_histogram() +
  ggtitle("Rate of Maternal Deaths from Abortions") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(p27)
figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
ggsave(filename = figure_file, plot=p27)

# explore outliers in maternal mortality from abortions
mydata %>% filter(rate_matdeaths_abs > 500)

```

P26: The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad and South Sudan that account for the highest rates of maternal mortality in this category.

P27: The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.