### 2024-03-28 code/analysis-code/statistical-analysis.qmd
Statistical Analysis Updates
I added another model with a combination of predictors and found more RMSE values to compare models. I found correlations between predictors and adjusted the multivariate model accordingly.

---
title: "Statistical Analysis"
author: "Taylor Glass and Arlyn Santiago"
date: "03/12/24"
output: html_document
---
## Load necessary packages
```{r}
#| message: false
library(dplyr) # for basic syntax
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) # for model building 
library(yardstick) # for finding performance metrics 
library(rsample) #for splitting the data into testing/training
```

## Load the cleaned data
```{r}
#path to data
data_location <- here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

## Data fitting and statistical analysis 
The outcome of interest in all of the following models is the percentage of unintended pregnancies associated with usage of short acting reversible contraception methods. The predictors used in our models include total number of women using short acting reversible methods standardized by population size, region, percent of women currently married, percent of women never married, total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy, and rate of maternal deaths due to abortion. We are considering adding another predictor using a political freedom index for each country.

### First Model Fit 
The first model attempts to predict the percentage of unintended pregnancies among women using short acting reversible methods of contraception 'pct_upreg_sarc' based on the total number of women using this method of contraception. We will use the version of the variable that has been standardized by population size, 'sarc_standard'. 
```{r}
# fit linear model using pct_upreg_sarc as outcome, sarc_standard as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit1 <- lm_mod %>% 
            fit(pct_upreg_sarc ~ sarc_standard, mydata) #estimate/train the model

#generate clean output with estimates and p-values
tidy(lmfit1)

# save results from fit into a data frame 
lmtable1 <- tidy(lmfit1)

# save fit results table  
table_file1 = here("results", "tables", "resulttable1.rds")
saveRDS(lmtable1, file = table_file1)
```
There appears to be a very strong relationship between total number of women using this birth control method and the outcome based on the coefficient estimate of 87.65998.

### Second Model Fit 
```{r}
# fit linear model using pct_upreg_sarc as outcome, region as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit2 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_larcster ~ region, mydata)

#generate clean output with estimates and p-values
tidy(lmfit2)

# save results from fit into a data frame 
lmtable2 <- tidy(lmfit2)

# save fit results table  
table_file2 = here("results", "tables", "resulttable2.rds")
saveRDS(lmtable2, file = table_file2)

```
This model shows that the Latin America/the Caribbean region has the strongest association with the outcome variable based on a coefficient estimate of 1.6218844 followed closely by the Asia region. The Oceania region has the weakest association with the outcome variable based on a coefficient estimate of 0.2512100.

### ANOVA for region variable
We want to compare the mean percent of unintended pregnancies among those using short acting reversible methods across the five regions: Asia, Africa, Europe, Latin America and the Caribbean, and Oceania. We will start with an ANOVA test to determine the relationship between region and outcome. We are more interested in pairwise comparisons between the regions, so we also conducted Tukey's Honestly Significant Difference test for pairwise comparisons within ANOVA. 
```{r}
# conduct an ANOVA test with the region variable
regionANOVA <- aov(pct_upreg_sarc ~ region, data = mydata)

# explore ANOVA results 
summary(regionANOVA)

# add Tukey's HSD test for pairwise comparisons
tukey_result <- TukeyHSD(regionANOVA)

# explore the pairwise comparisons
tidy(tukey_result)
```
ANOVA results show that the region variable is statistically significantly associated with the outcome variable given the p-value less than 0.001. The Tukey HSD test revealed the largest differences in the outcome between the Oceania region and Latin America/the Caribbean, Europe, and Asia. The relationships between Oceania and Europe and Latin America/the Caribbean are statistically significant with p-values less than 0.0001. Oceania is clearly experiencing the lowest percentage of unintended pregnancies among short acting reversible method users.  The lowest difference in outcome between regions is between Asia and Africa, but this is not a statistically significant association. Two statistically significant differences in the outcome exist between Latin America/the Caribbean and Europe with Africa, which shows that Africa is experiencing higher levels of the outcome.

### Third Model Fit 
```{r}
# fit linear model using pct_upreg_sarc as outcome, pct_currentlymarried as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit3 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_sarc ~ pct_currentlymarried, mydata)

#generate clean output with estimates and p-values
tidy(lmfit3)

# save results from fit into a data frame 
lmtable3 <- tidy(lmfit3)

# save fit results table  
table_file3 = here("results", "tables", "resulttable3.rds")
saveRDS(lmtable3, file = table_file3)
```
This model shows there is a negative relationship between the number of women currently married and the outcome. As the percentage of currently married women increases by 1, the percentage of unintended pregnancies decreases by a factor of 0.4181571.

# Fourth Model Fit 
```{r}
# fit linear model using pct_upreg_sarc as outcome, pct_nevermarried as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit4 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_sarc ~ pct_nevermarried, mydata)

#generate clean output with estimates and p-values
tidy(lmfit4)

# save results from fit into a data frame 
lmtable4 <- tidy(lmfit4)

# save fit results table  
table_file4 = here("results", "tables", "resulttable4.rds")
saveRDS(lmtable4, file = table_file4)
```
This model shows a positive relationship between the percentage of women never married and the outcome. As the percentage of women never married increases by 1, the percentage of unintended pregnancies increases by a factor of 0.2474218.

# Fifth Model Fit 
```{r}
# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit5 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_sarc ~ rate_matdeaths, mydata)

#generate clean output with estimates and p-values
tidy(lmfit5)

# save results from fit into a data frame 
lmtable5 <- tidy(lmfit5)

# save fit results table  
table_file5 = here("results", "tables", "resulttable5.rds")
saveRDS(lmtable5, file = table_file5)
```
This model shows that total rate of maternal deaths is slightly negatively correlated with the outcome based on a small coefficient estimate of -0.01482568.

# Sixth Model Fit 
```{r}
# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths_upreg as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit6 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_sarc ~ rate_matdeaths_upreg, mydata)

#generate clean output with estimates and p-values
tidy(lmfit6)

# save results from fit into a data frame 
lmtable6 <- tidy(lmfit6)

# save fit results table  
table_file6 = here("results", "tables", "resulttable6.rds")
saveRDS(lmtable6, file = table_file6)
```
This model shows that total rate of maternal deaths due to unintended pregnancies has an even weaker association with the outcome based on a small coefficient estimate of -0.003158542.

# Seventh Model Fit 
```{r}
# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths_abs as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit7 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_sarc ~ rate_matdeaths_abs, mydata)

#generate clean output with estimates and p-values
tidy(lmfit7)

# save results from fit into a data frame 
lmtable7 <- tidy(lmfit7)

# save fit results table  
table_file7 = here("results", "tables", "resulttable7.rds")
saveRDS(lmtable7, file = table_file7)
```
This model shows that total rate of maternal deaths due to abortion has the strongest association with the outcome out of the three maternal mortality rates included based on a coefficient estimate of -0.033353527.

### Assessing correlation between predictors
There are some variables that are intuitively correlated, which means they should not be included in the same model. We want to explore those correlations before building our multivariate linear regression model. 
```{r}
# find correlation coefficients 
cor(mydata$pct_currentlymarried, mydata$pct_nevermarried)
cor(mydata$rate_matdeaths, mydata$rate_matdeaths_upreg)
cor(mydata$rate_matdeaths, mydata$rate_matdeaths_abs)
cor(mydata$rate_matdeaths_upreg, mydata$rate_matdeaths_abs)
```
The percent of currently married women and the percent of never married women have a correlation coefficient of -0.8588, so these two variables are highly correlated. The total rate of maternal deaths and rate of maternal deaths due to unintended pregnancies is the most highly correlated pair with a coefficient of 0.8904. The correlation between total rate of maternal deaths and rate of maternal deaths due to abortion is slightly lower with a coefficient of 0.8023. The two variables with the least correlation out of these four pairs are rate of maternal deaths due to unintended pregnancies and rate of maternal deaths due to abortion with a coefficient of 0.6486. Based on these associations, we will only include the percent of currently married women in the multivariate model because it has a slightly larger effect on the outcome with a more significant p-value in its single predictor model compared to the percent of never married women predictor. We will include both the rate of maternal deaths due to unintended pregnancies and rate of maternal deaths due to abortion beccause the total rate of maternal deaths is so highly correlated with each of these variables.

### Eighth Model Fit 
Based on the exploration of correlation between predictors, we include 5 predictors in the multiple linear regression model to determine how the associations interact. 
```{r}
# fit linear model using pct_upreg_sarc as outcome with all 7 predictors
lm_mod <- linear_reg() #specify the type of model
lmfit8 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg + rate_matdeaths_abs, mydata)

#generate clean output with estimates and p-values
tidy(lmfit8)

# save results from fit into a data frame 
lmtable8 <- tidy(lmfit8)

# save fit results table  
table_file8 = here("results", "tables", "resulttable8.rds")
saveRDS(lmtable8, file = table_file8)
```
This model shows the strongest predictor is the number of women using short acting reversible contraception with a huge coefficient of 85.6433. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. Both of the marriage status predictors are associated with a decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome.

### Nineth model fit 
It is important to create the null model for future assessment of model performance, which just calculates the mean outcome without any predictors present. 
```{r}
# create a null model
null_model <- null_model(mode = "regression") %>% 
    set_engine("parsnip") %>%
    fit(pct_upreg_sarc ~ 1, data = mydata)

#generate clean output with estimates and p-values
tidy(null_model)

# save results from fit into a data frame 
nulltable <- tidy(null_model)

# save fit results table  
table_file9 = here("results", "tables", "resulttable9.rds")
saveRDS(nulltable, file = table_file9)
```
We also want to observe models with different combinations of predictors to determine if there is any evidence of interaction or effect measure modification. Multivariate analysis allows for the assessment of the association between each predictor and the outcome while adjusting for the effects of other variables.

### Tenth Model fit 
We will create a model with all 3 rate of maternal mortality predictors to compare to the three models with each predictor on its own to further explore the correlation between these variables. 
```{r}
# fit linear model using pct_upreg_sarc as outcome with rate of maternal mortality predictors
lm_mod <- linear_reg() #specify the type of model
lmfit10 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_sarc ~ rate_matdeaths + rate_matdeaths_upreg + rate_matdeaths_abs, mydata)

#generate clean output with estimates and p-values
tidy(lmfit10)

# save results from fit into a data frame 
lmtable10 <- tidy(lmfit10)

# save fit results table  
table_file10 = here("results", "tables", "resulttable10.rds")
saveRDS(lmtable10, file = table_file10)
```
It appears that there is an association between these three variables because the direction of the association between rate of maternal deaths and outcome has flipped. This model shows that rate of maternal deaths is associated with an increase in the outcome by a factor of 0.011267, while the single predictor model showed a slightly negatively correlation with the outcome based on a small coefficient estimate of -0.01482568. The other two variables retain their negative association with the outcome variable with similar coefficient measurements.

## Assessing Model Performance
It is important to consider appropriate metrics to measure model performance. We will use RMSE at this point to compare 3 models: one with the number of women using short acting reversible contraception predictor, one using the percent never married predictor, and one using the rate of maternal deaths due to abortion predictor. We chose these predictors because they have the strongest associations with the outcomes based on the single linear regression models shown above. 
```{r}
# set seed for reproducibility 
set.seed(2468)

# create predictions using the sarc_standard predictor model
poppred <- predict(lmfit1, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
popdata <- bind_cols(poppred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse1 <- rmse(popdata, truth = pct_upreg_sarc, estimate = .pred)

# create predictions using the pct_curmarried predictor model
marriedpred <- predict(lmfit3, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
marrieddata <- bind_cols(marriedpred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse2 <- rmse(marrieddata, truth = pct_upreg_sarc, estimate = .pred)

# create predictions using the rate of maternal deaths due to abortion predictor model
matdeaths_abspred <- predict(lmfit7, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
matdeath_absdata <- bind_cols(matdeaths_abspred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse3 <- rmse(matdeath_absdata, truth = pct_upreg_sarc, estimate = .pred)

# create predictions using the null model
nullpred <- predict(lmfit9, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
nulldata <- bind_cols(nullpred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse4 <- rmse(nulldata, truth = pct_upreg_sarc, estimate = .pred)

# compare the metrics 
print(rmse1)
print(rmse2)
print(rmse3)
print(rmse4)
```

The RMSE metric favors the model with the total number of women using short acting reversible contraception methods based on a low RMSE value of 5.143319. The model with the percentage of women currently married performs worse with an RMSE value of 10.17591. The same conclusion can be drawn for the model with the rate of maternal mortality due to abortions predictor only, given the RMSE value of 10.14128. The null model has an RMSE value of 9.842267, so all models should have an RMSE lower than this to be considered useful.

It is important to determine if our model with the region predictor performs well because the main question to be answered by this analysis depends on this predictor. Since region is a categorical variable, we cannot evaluate the RMSE performance metric.  
```{r}
# set seed for reproducibility 
set.seed(2468)

# create predictions using the sarc_standard predictor model
regionpred <- predict(lmfit2, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
regiondata <- bind_cols(regionpred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse5 <- rmse(regiondata, truth = pct_upreg_sarc, estimate = .pred)
print(rmse5)
```

The results of this RMSE value are disappointing because 18.11857 is much larger than the null model's RMSE of 9.842267. However, this issue could possibly be explained by the fact that Africa has the most significant differences in our outcome compared to the other regions. We will still include this predictor in our multivariate regression model based on the significant Tukey conclusions and the large differences in outcome displayed in our exploratory data analysis results.

After examining the relationship between the rate of maternal mortality variables, we want to determine which model performs the best. We have already found the RMSE value for the rate of maternal deaths due to abortion, which is stored as `rmse3`. We need to find the RMSE values for the other two single predictor models and the model with all 3 predictors.  
```{r}
# create predictions using the rate of maternal deaths predictor model
matdeathsallpred <- predict(lmfit5, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
matdeathalldata <- bind_cols(matdeathsallpred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse6 <- rmse(matdeathalldata, truth = pct_upreg_sarc, estimate = .pred)
print(rmse6)

# create predictions using the rate of maternal deaths from unintended pregnancies predictor model
matdeaths_upregpred <- predict(lmfit6, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
matdeath_upregdata <- bind_cols(matdeaths_upregpred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse7 <- rmse(matdeath_upregdata, truth = pct_upreg_sarc, estimate = .pred)
print(rmse7)

# create predictions using model with all three rate predictors
totalratepred <- predict(lmfit10, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
totalratedata <- bind_cols(totalratepred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse8 <- rmse(totalratedata, truth = pct_upreg_sarc, estimate = .pred)
print(rmse8)

# display RMSE values for each model
print(rmse6) #rate of maternal mortality predictor only
print(rmse7) #rate of maternal mortality due to unintended pregnancy predictor only
print(rmse3) #rate of maternal mortality due to abortion predictor only
print(rmse8) #all 3 rates of maternal mortality predictors together 
```
The RMSE values shows that the three single predictor models perform worse than the null model, which is not ideal. The model with all 3 maternal morality rate predictors performs exactly the same as the null model, which makes sense because the coefficients estimating the effect of each predictor on the outcome are so small.

We want to see how the model with all 5 predictors performs in comparison to the single predictor models. 
```{r}
# set seed for reproducibility 
set.seed(2468)

# create predictions using the multiple linear regression model
allpred <- predict(lmfit8, new_data = mydata) %>% 
                  select(.pred)

# create a data frame with the predictions and true values 
alldata <- bind_cols(allpred, mydata$pct_upreg_sarc) %>%
            rename(pct_upreg_sarc = "...2")

# find RMSE to determine model fit
rmse9 <- rmse(alldata, truth = pct_upreg_sarc, estimate = .pred)
print(rmse9)
```
As expected, this model performs better than the single predictor models with the lowest RMSE score of 4.209.

## Cross-Validation to Assess Model Performance
While the RMSE performance is a good place to start for evaluating model performance, we need to determine how well the model performs on data it has not seen before to determine its usefulness. We will use the cross-validation process to do this. The first step is to split the data into testing and training portions. 
```{r}
# set seed for reproducibility
set.seed(2468)

# put 75% into the training data
datasplit <- initial_split(mydata, prop = 3/4)

# create data frames for the two sets
train_data <- training(datasplit)
test_data <- testing(datasplit)
```

We will only cross-validate the multivariate regression model `lmfit8` because it performs the best in terms of RMSE. 
```{r}

```