### 2024-04-05 products/manuscript/supplement/Supplementary-Material.qmd
Part 4 Updates
Added method descriptions in the supplemental file and wrote instructions for reproducibility. Updated manuscript with code processing summary and CV/bootstrapping explanations.

---
title: "Supplement to Global Distribution of Unintended Pregnancies and Contraceptive Usage Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis_template_references.bib
csl: ../../assets/vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
#load necessary packages
library(here) #for data loading/saving
library(dplyr) # for basic piping functions
library(skimr) # for summarizing data
library(ggplot2) # for visualizing data
library(gtsummary) # for creating tables
```

# Overview
This file provides extra information that is not included in our manuscript or the main processing and exploratory data analysis files. This information provides guidance on our thought processes behind the scenes and explains why we chose a single outcome to study.

# Code and file information

Reproducing this project requires R, RStudio, Microsoft word, and LaTex. The README files in each folder contain descriptions of the documents located in that folder to help guide users through the reproducible processes.

Begin with the processing-file.qmd in the code > processing-code folder. This file loads the original dataset from the Guttmacher Institute's Adding It Up project, loads the codebook for the dataset, and glimpses at the data. The processing-code.qmd file shows how we modified existing variable, created new variables, and saved the processed data. The processed data is saved in the data > processed-data folder as processeddata.rds.

Visit the eda.qmd file in the code > eda-code folder to find code for the exploratory data analysis process. This code generates all the exploratory figures presented in the Exploratory/descriptive analysis results in the manuscript.qmd file as well as additional figures for the outcome not selected for the final analysis as shown later in this supplementary material.  All exploratory figures are saved in the results > figures > exploratory-figures folder.

Continue to the statistical-analysis.qmd file in the code > analysis-code folder to find code for the statistical analysis and model building portion of the study. All model results tables are stored in the results > tables folder. All model figures examining model performance are stored in the results > figures > final-figures folder.

View the manuscript in the products > manuscript folder to find a summary of the study along with the most important findings and final conclusions. This supplemental material is stored in the products > manuscript > supplement folder.

{{< pagebreak >}}

# Additional Method Details

We only briefly mentioned that we used cross-validation, bootstrapping, and LASSO regression to test our models in the methods section of our manuscript. Cross-validation was chosen as a method for determining how well the model performs on data it has not seen before because that is the main point of building models. Our original models, found in resulttables1-10 in the results > tables folder, are created using all of the data. The cross-validation process built models with 75% of the data that was designated as the training data, and the remaining 25% was used as testing data. Our dataset is small with only 132 observations, so the insights provided here are not robust. However, we still want to know how well these variables predict our outcome to determine how meaningful each predictor is, so we completed cross-validation anyways to answer this question. We also used this approach due to the risk of overfitting in our model based on interaction between variables. Plotting the observed and predicted values for our model allowed us to look for systematic deviations from the expected relationship visualized with a 45 degree line as shown in @supplement-1. Since we did conclude that the model overfits the data based on unhealthy deviations of the testing data from the 45 degree line, our use of cross-validation is appropriate and useful. Lastly, this approach is an easily interpretable way to increase confidence in the performance of our models based on metrics because both proceses lead to the same conclusions. 
```{r}
#| label: supplement-1
#| fig-cap: "Evaluating Final Model Performance on Test Data"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
```

Bootstrapping was chosen to determine variability and uncertainty in the data. This process is similar to cross-validation, but the resampling process simulates a new dataset by performing sampling with replacement to obtain the same sample size as the original dataset. This is crucial for our small dataset of 132 observations, which is why we chose to bootstrap with 100 resamples. We are still limited by our small sample size for this approach, but it is worth completing for the purpose of investigating uncertainty. The model estimates parameters for each of the 100 samples, which creates a distribution of parameter estimates that can be used to create confidence intervals. We chose to create standard 95% confidence intervals based on convention, which can be seen in @supplement-2. 
```{r}
#| label: supplement-2
#| fig-cap: "Evaluation of Uncertainty in Full Model Predictions"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "bootstrappred.png"))
```

LASSO regression is important in this study because we need to manage the bias-variance tradeoff since we have evidence that our model overfits the data. By applying a penalty parameter to the cost function, LASSO regression shrinks the coefficients of the parameter estimates, so these estimates will always be smaller than the ones provided by the original linear model. As the penalty parameter increases, a larger penalty term is added to the SSR, which increases the value of the cost function. As the penalty increases, the RMSE value increases, and the performance of the model is considered worse and worse comparatively. We chose a parameter of 0.1 first because that is standard for LASSO regression, but we also completed a parameter tuning process to validate that choice. As shown in @supplement-3, the RMSE performance metric increases exponentially when the penalty parameter increases past 1, which validates our previous selection of 0.1 for the penalty parameter because that will produce the lowest RMSE possible. Since coefficients can be reduced all the way to zero, this analysis completes feature selection as well. Shrinking the coefficients also helps combat colinearity in variables, which we established between the percent of currently married women and percent of never married women, and the three rates of maternal mortality, respectively. LASSO models are also easier to interpet, which increases the validity of our findings.
```{r}
#| label: supplement-3
#| fig-cap: "Parameter Tuning for LASSO Regression"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
```

# Additional data cleaning explanation
The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable. We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained. This variable is only 82% complete, but the individual percent of unintended pregnancy variables correponding with different types of birth control methods are each 100% complete. Since the missingness of this data is not well explained except by the hypothesis that this variable was found using methods outside of the AIU protocol, we decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information. We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence. 
```{r}
#| label: tbl-summarytable2
#| tbl-cap: "Summary table to discover missingness"
#| echo: FALSE
resulttable2 <- readRDS("../../results/tables/summarytable.rds")
kable(resulttable2)
```

{{< pagebreak >}}

# Additional results

Readers will be able to find our analysis of the different contraception methods that we did not select for our main outcome through figures we've made for the `pct_upreg_larcster`, `pct_upreg_trad`, and `pct_upreg_nouse` variables in this section.

```{r}
# explore the percentage of unintended pregnancies related to long acting reversible methods in each country
p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + 
  geom_histogram() +
  ggtitle("Percent Unintended Pregnancy Using Long Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(p2)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_distribution.png")
ggsave(filename = figure_file, plot=p2)

# explore the percentage of unintended pregnancies related to traditional methods in each country
p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + 
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using Traditional Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(p4)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_distribution.png")
ggsave(filename = figure_file, plot=p4)

# explore the percentage of unintended pregnancies related to no birth control usage in each country
p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) +
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using No Birth Control") + xlab("Percentage") + ylab("Number of countries")
plot(p5)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_nouse_distribution.png")
ggsave(filename = figure_file, plot=p5) 
```

P2: This histogram is highly skewed right, which shows that most countries experience a very low percentage of unintended pregnancies related to long acting reversible and sterilization methods.

P4: This histogram shows that about half of the countries reported less than 10% percent of unintended pregnancies associated with using traditional birth control methods

P5: This histogram shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.

The following set of plots revealed large outliers in the 'total_larcster`,'total_trad', and `total_none` variables, which prevented the graphs from being very useful. The histograms look like most observations for each of these four variables are 0, but that is not possible because the summary table showed that the minimum value for all four of these variables is greater than 0. This distortion is caused by the two large outliers in the data set.

```{r}
# explore the total number of people using long acting reversible methods in each country
p6 <- mydata %>% ggplot(aes(x=total_larcster)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Long Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p6)
figure_file = here("results", "figures", "exploratory-figures", "total_larcster_distribution.png")
ggsave(filename = figure_file, plot=p6)

# explore the total number of people using traditional methods in each country
p8 <- mydata %>% ggplot(aes(x=total_trad)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Traditional Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p8)
figure_file = here("results", "figures", "exploratory-figures", "total_trad_distribution.png")
ggsave(filename = figure_file, plot=p8)

# explore the total number of people not using birth control in each country
p9 <- mydata %>% ggplot(aes(x=total_none)) + 
  geom_histogram() + 
  ggtitle("Total Usage of No Birth Control Method per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p9)
figure_file = here("results", "figures", "exploratory-figures", "total_none_distribution.png")
ggsave(filename = figure_file, plot=p9)

```

Plot 10 shows `pct_upreg_larcster` vs `larcster_standard` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.

```{r}
# explore percentage of unintended pregnancies compared to standardized count of women using long-acting reversible and sterilization methods
p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p10)
figure_file = here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png")
ggsave(filename = figure_file, plot=p10) 
```

P10: As the standardized count of women using LARCs and sterilization methods increases, the percentage of unintended pregnancies tends to increase. There is minimal variation around the linear equation, excluding the two largest percentages accounted for by India and China.

Plot 12 shows `pct_upreg_trad` versus `trad_standard` 
```{r}
# explore percentage of unintended pregnancies compared to standardized count of women using traditional methods
p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p12)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png")
ggsave(filename = figure_file, plot=p12) 
```

P12: As percentage of unintended pregnancies associated with traditional methods increases, the total number of women using this method increases. There is minimal variation in the trend when percentage of unintended pregnancies is less than 8%, but there is more variation when the percentage is greater than 10%.

Plot 13 shows `pct_upreg_nouse` versus `none_standard`
```{r}
# explore percentage of unintended pregnancies compared to standardized count of women not using no method
p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p13)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png")
ggsave(filename = figure_file, plot=p13) 
```

P13: The same expected positive trend is observed for women who are not using a birth control method, but there is considerable variation in this graph across all percentages of unintended pregnancies. Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.

We are also interested in how these relationships between percent of unintended pregnancies using each birth control method and the number of people using each method varies across region. We adjusted the original scatter plots created above by coloring based on region.
```{r}
# explore percent unintended pregnancies using long acting reversible and sterilziation methods by region
p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + 
  geom_point() + geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p14)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png")
ggsave(filename = figure_file, plot=p14)

# explore percent unintended pregnancies using traditional methods by region
p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') +  
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p16)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png")
ggsave(filename = figure_file, plot=p16)

# explore percent unintended pregnancies using no method by region
p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') +   
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p17)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png")
ggsave(filename = figure_file, plot=p17)

```

P14: Among women using long acting reversible and sterilization methods, Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.

P16: Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.

P17: This graph is much different than the other three, which is to be expected at this point. Oceania has the highest percentage of unintended pregnancies when women are not using birth control. The slope of the relationship for each region appears to be similar in this graph with the exception of Oceania having a steeper slope. There are considerably more outliers in this graph with African countries

Our next group of plots will be violin plots to explore the usage level of each birth control method, a categorical variable, associated with percentage of unintended pregnancies use for each birth control method, a numerical variable.
```{r}
# explore long acting reversible and sterilization method usage compared to percent of unintended pregnancies
p18 <- mydata %>% ggplot(aes(x = larcster_usage, y = pct_upreg_larcster)) +
  geom_violin() +
  ggtitle("Usage of LARCs and Sterilization vs Percent of Unintended Pregnancies") +
  labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
plot(p18)
figure_file = here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p18)

# explore traditional method usage compared to percent of unintended pregnancies
p20 <- mydata %>% ggplot(aes(x = trad_usage, y = pct_upreg_trad)) +
  geom_violin() + 
  ggtitle("Usage of Traditional Method vs Percent of Unintended Pregnancies") +
  labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
plot(p20)
figure_file = here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p20)

# explore no method usage compared to percent of unintended pregnancies
p21 <- mydata %>% ggplot(aes(x = none_usage, y = pct_upreg_nouse)) +
  geom_violin() +  
  ggtitle("No Birth Control Usage vs Percent of Unintended Pregnancies") +
  labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
plot(p21)
figure_file = here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p21)

```

P18: The distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods. High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.

P20: Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.

P21: While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.

We explored the distribution of the countries by five categorical variables: `region`, `larcster_usage`, `sarc_usage`, `trad_usage`, and `none_usage` which is displayed in @tbl-summarytable1. These variables were not used in the main final analysi of the paper. We found that most countries are located in Africa followed by Asia and Latin America/the Caribbean. Only 20 countries have high usage of long acting reversible and sterilization methods compared to 33 countries with high usage of short acting reversible and traditional methods. It is interesting that the distributions for short acting reversible method, traditional method, and no method usages are the same. 
```{r}
#| label: tbl-summarytable1
#| tbl-cap: "Summary table for categorical variables"
#| echo: FALSE
resulttable1 <- readRDS("../../results/tables/summarytablecat.rds")
kable(resulttable1)
```

# Additional data cleaning explanation
The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable. We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained. We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information. We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence. 
```{r}
#| label: tbl-summarytable2
#| tbl-cap: "Summary table to discover missingness"
#| echo: FALSE
resulttable2 <- readRDS("../../results/tables/summarytable.rds")
kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

{{< pagebreak >}}

# References